---
title: "Initial Korean covid 19"
author: "Tiffany Alvarado"
date: "FAll 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Methods 

Since is sex is catagorical and result is catagorical we use the cat_cat helper file.   

$H_0$ dying is not independent of sex.  

$H_A$ Dying is independent of sex.  

## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
This shows there are more dead men than there are alive.
  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

among the people who died there were 64% of the men who died and 36% of women who died.  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi spread test was 0.01 so we reject the null hypothesis. we conclude that dying is dependent of sex. therefor we are 2.3 times more likely to die from covid in males than females in korea.  

## Conclusions

In conclusion we reject the null hypothesis that dying is not independent of sex. 64% of males died versus the 36% of women who died. In the fisher test results it shows that we are 2.3 times more likely to die from covid in males than females.
